### 2025-07-22 machine-learning.qmd
20250722 - fix error after new gpboost package version

A learning rate of 1 is very high for boosting.
Even if a learning rate of 1 did well in tuning, we use a lower learning rate (0.1) to avoid overfitting.
We also added some light regularization (`lambda_l2`) for better generalization.
We also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^5 (32).
We set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 10.

@@

params_qb <- list(
  learning_rate = 0.1,
  max_depth = 5,
  min_data_in_leaf = 10,
  lambda_l2 = 1,
  num_leaves = 2^5,
  num_threads = num_cores)

nrounds_qb <- 123 # identify optimal number of trees through iteration and cross-validation

>>>>

We use the optimal parameters identified during tuning.
We use a low learning rate (0.1) to avoid overfitting.
We set some light regularization (`lambda_l2`) for better generalization.
We also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^6 (64)—though the number will not reach greater than $2^\text{max_depth} = 2^5 = 32$.
We set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 100.

@@

params_qb <- list(
  learning_rate = 0.1,
  max_depth = 5,
  min_data_in_leaf = 100,
  lambda_l2 = 5,
  num_leaves = 2^6,
  num_threads = num_cores)

nrounds_qb <- 84 # identify optimal number of trees through iteration and cross-validation
